---
title: "earlywarnings vignette"
author: "Vasilis Dakos and Leo Lahti"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
    theme: united
    highlight: pygments
---

<!--
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{microbiome tutorial}
  %\usepackage[utf8]{inputenc}
-->



earlywarnings R package
===========

## Installation

### Installing and loading the release version

<!-- Note: if dependencies are missing you may wish to run the [../installation.script](installation.script) first -->

```{r install, eval=FALSE}
install.packages("earlywarnings")
```

### Installing and loading the experimental development version

```{r install2, eval=FALSE}
library(devtools)
install_github("earlywarningtoolbox/earlywarnings-R/earlywarnings")
```

### Clone the repository & install locally:

Run on command line:
<pre><code>git clone git@github.com:earlywarningtoolbox/earlywarnings-R.git
./build.sh
</pre></code>

### Loading the package

```{r loading, eval=TRUE}
library(earlywarnings)  
```


## Potential analysis

 Potential analysis, used for instance in [Hirota et al. Science, 2011, 334, 232-235.]<!--(https://www.science.org/doi/full/10.1126/science.1210657) -->

```{r movpotential, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
# Create simulated example data
X <- c(rnorm(1000, mean = 0), rnorm(1000, mean = -2), 
 	           rnorm(1000, mean = 2))
param <- seq(0,5,length=3000) 

# Run potential analysis
res <- movpotential_ews(X, param)

# Visualize
p <- PlotPotential(res$res, title = '', 
	       	   xlab.text = '', ylab.text = '', 
		   cutoff = 0.5, plot.contours = TRUE, binwidth = 0.2)
print(p)
```


### Licensing and Citations

This work can be freely used, modified and distributed under the 
[Two-clause (Free)BSD license](https://en.wikipedia.org/wiki/BSD\_licenses).

Kindly cite the work as 'Vasilis Dakos and Leo Lahti (2014). earlywarnings R package. URL: https://github.com/earlywarningtoolbox/earlywarnings-R'.


```{r citation}
citation("earlywarnings")
```

### Session info

This vignette was created with

```{r sessioninfo}
sessionInfo()
```




